---
title:  'Computational Musicology'
author: 'John Ashley Burgoyne'
date:   'February--March 2020'
output: 
    flexdashboard::flex_dashboard:
        storyboard: true
        theme: flatly
    github_document
---

```{r setup}
# In order to use these packages, we need to install flexdashboard, plotly, and Cairo.
library(tidyverse)
library(plotly)
library(spotifyr)
library(Cairo)
source('spotify.R')
```

Instructions:
```{r}
### Three hashtags create a panel block. Use a sentence that describes the contents (so no introductions et cetera)
```

### The Musically shapeshifting genius that is Miles Davis 

For my corpus I have decided to investigate the jazz trumpet player Miles Davis. Davis is known within the jazz and global community as a constant innovator and key figure of jazz music. His method for innovation was to constantly attract upcoming talents and stars within the upcoming current in jazz. He then instructed these youngsters with his own vision of jazz and this was reciprocated by the youngsters how provided him with their own new views. 

In my corpus I will try to analyse Davis’ music by answering some of the following questions:

-	How did Davis’ music change over time, what are the constants in his music and what are the variables?

-	How did Davis’ music reflect other albums of each specific time period that he was adapting to?


The following phases of Miles Davis’ oeuvre could be examined

-	Early bebop/Charlie Parker phase: 1945-1950

-	Hard Bop phase: 1950-1955

-	Miles Davis Cool Phase: 1955-1963

-	Miles Davis Acoustic Erratic: 1963-1968

-	Miles Davis electric/funk fusion: 1968-1975

-	Miles Davis Hiatus: 1975-1980

- Miles Davis final years: 1980-1992
```{r echo=FALSE, message=FALSE, warning=FALSE}
miles_davis <- get_artist_audio_features("Miles Davis")
#Miles_Parker(1944-1950)

miles40ies <- get_playlist_audio_features("Wpleiter", "1v8oL4Clr0LliAh6HNK4C2")
miles40ies_cut <- as.data.frame(miles40ies[,c(6:16)])
miles40ies_cut$Time_Period <- "1944-1950"
#1950-1955
Miles_early_50s <- get_playlist_audio_features("Wpleiter", "3lv2u17pbcXIoNZZtP8wZ5")
milesE50ies <- as.data.frame(c(Miles_early_50s[,c(6:16)]))
milesE50ies$Time_Period <- "1951-1954"

#first quintet and 1955-1962
milesmodal <- subset(miles_davis, album_release_year > 1954 & album_release_year < 1963)
milesmodal_short <- as.data.frame(milesmodal[,c(9:19)])
milesmodal_short$Time_Period <- "1955-1962"


#second quintet 1963-1968
miles_herbie <- subset(miles_davis, album_release_year > 1963 & album_release_year < 1969)
miles_herbie_short <- as.data.frame(miles_herbie[,c(9:19)])
miles_herbie_short$Time_Period <- "1963-1968"

#Milesfusion:
miles_fusion <- subset(miles_davis, album_release_year > 1969 & album_release_year < 1975)
miles_fusionshort <- as.data.frame(miles_fusion[,c(9:19)])
miles_fusionshort$Time_Period <- "1969-1975"

#Miles_hiatus
miles_hiatus <- subset(miles_davis, album_release_year > 1976 & album_release_year < 1979)
miles_hiatus_short <- as.data.frame(miles_hiatus[,c(9:19)])
miles_hiatus_short$Time_Period <- "1976 - 1979"

#80ies_todeath
miles_death <- subset(miles_davis, album_release_year > 1979 & album_release_year < 1992)
miles_death_short <- as.data.frame(miles_death[,c(9:19)])
miles_death_short$Time_Period <- "1980 - 1992"

miles_long <- rbind(miles40ies_cut, milesE50ies, milesmodal_short, miles_herbie_short, miles_fusionshort, miles_hiatus_short, miles_death_short)
miles_wide <- miles_long %>%
  gather("danceability", "energy", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", key = Feature, value = Parameter)

```


***

Placeholder for a picture or collage of miles'albums or paintings



### Miles transformations are visible in data from Spotify

```{r echo=FALSE, message=FALSE, warning=FALSE}

#Slopegraph is not in my gg-package for some reason, trying something else:
if (!require("remotes")) {
    install.packages("remotes")
}
remotes::install_github("leeper/slopegraph")

library("slopegraph")

#Creating a unique matrix to fulfill the conditions of ggplot
Miles_slope <- matrix(0, 7, 7)
colnames(Miles_slope) <- unique(miles_wide$Time_Period)
rownames(Miles_slope) <- names(miles_death_short[,c(1,2,6,7,8,9,10)])


#40ies
Miles_slope[1,1] <- mean(miles40ies_cut[,1])
Miles_slope[2,1] <- mean(miles40ies_cut[,2])
Miles_slope[3,1] <- mean(miles40ies_cut[,6])
Miles_slope[4,1] <- mean(miles40ies_cut[,7])
Miles_slope[5,1] <- mean(miles40ies_cut[,8])
Miles_slope[6,1] <- mean(miles40ies_cut[,9])
Miles_slope[7,1] <- mean(miles40ies_cut[,10])



#second
Miles_slope[1,2] <- mean(milesE50ies[,1])
Miles_slope[2,2] <- mean(milesE50ies[,2])
Miles_slope[3,2] <- mean(milesE50ies[,6])
Miles_slope[4,2] <- mean(milesE50ies[,7])
Miles_slope[5,2] <- mean(milesE50ies[,8])
Miles_slope[6,2] <- mean(milesE50ies[,9])
Miles_slope[7,2] <- mean(milesE50ies[,10])


#Third
Miles_slope[1,3] <- mean(milesmodal_short[,1])
Miles_slope[2,3] <- mean(milesmodal_short[,2])
Miles_slope[3,3] <- mean(milesmodal_short[,6])
Miles_slope[4,3] <- mean(milesmodal_short[,7])
Miles_slope[5,3] <- mean(milesmodal_short[,8])
Miles_slope[6,3] <- mean(milesmodal_short[,9])
Miles_slope[7,3] <- mean(milesmodal_short[,10])


#4
Miles_slope[1,4] <- mean(miles_herbie_short[,1])
Miles_slope[2,4] <- mean(miles_herbie_short[,2])
Miles_slope[3,4] <- mean(miles_herbie_short[,6])
Miles_slope[4,4] <- mean(miles_herbie_short[,7])
Miles_slope[5,4] <- mean(miles_herbie_short[,8])
Miles_slope[6,4] <- mean(miles_herbie_short[,9])
Miles_slope[7,4] <- mean(miles_herbie_short[,10])

#5
Miles_slope[1,5] <- mean(miles_fusionshort[,1])
Miles_slope[2,5] <- mean(miles_fusionshort[,2])
Miles_slope[3,5] <- mean(miles_fusionshort[,6])
Miles_slope[4,5] <- mean(miles_fusionshort[,7])
Miles_slope[5,5] <- mean(miles_fusionshort[,8])
Miles_slope[6,5] <- mean(miles_fusionshort[,9])
Miles_slope[7,5] <- mean(miles_fusionshort[,10])

#6
Miles_slope[1,6] <- mean(miles_hiatus_short[,1])
Miles_slope[2,6] <- mean(miles_hiatus_short[,2])
Miles_slope[3,6] <- mean(miles_hiatus_short[,6])
Miles_slope[4,6] <- mean(miles_hiatus_short[,7])
Miles_slope[5,6] <- mean(miles_hiatus_short[,8])
Miles_slope[6,6] <- mean(miles_hiatus_short[,9])
Miles_slope[7,6] <- mean(miles_hiatus_short[,10])

#7
Miles_slope[1,7] <- mean(miles_death_short[,1])
Miles_slope[2,7] <- mean(miles_death_short[,2])
Miles_slope[3,7] <- mean(miles_death_short[,6])
Miles_slope[4,7] <- mean(miles_death_short[,7])
Miles_slope[5,7] <- mean(miles_death_short[,8])
Miles_slope[6,7] <- mean(miles_death_short[,9])
Miles_slope[7,7] <- mean(miles_death_short[,10])


Miles_slope <- as.data.frame(Miles_slope)

data <- Miles_slope
xlabels <- names(data)

palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")
ggslopegraph(Miles_slope, offset.x = 0.06,xlim = c(-0.5,7.9),
  decimals = 2, col.lab = palette, col.lines = palette, col.num = palette, xlabels = c("'45-'50", "'51-'55", "'55-'63", "'63-'68", "'68-'75", "'75, '80", "'80-'92")) + 
  theme(panel.grid =element_blank(), panel.background = element_blank())

```















### Miles in terms of Valence, Danceability and energy
```{r echo=FALSE, message=FALSE, warning=FALSE}

# miles_plot <- ggplot(miles_decades, aes(x = danceability, y = energy, col = Time_Period)) + 
#   geom_jitter()
ggplot(miles_long, aes(x = danceability, y = energy, col = valence)) + 
  geom_point(size = 3, alpha = 0.6) +
  facet_wrap(. ~ Time_Period)

```


